The Impact of Past Behavior Normality on Regret: A Replication Study

### 2024-02-14 outputs/paper/paper.qmd
update

Regret is a pervasive and intense negative emotion, plays a pivotal role in shaping individuals' daily experiences. The recognition that other decisions or circumstances could have led to more favorable outcomes triggers regret (Byrne, 2016; Roese & Epstude, 2017; Kahneman & Tversky, 1982). Influential work by Kahneman and Miller (1986) in norm theory has highlighted the connection between abnormal behavior, heightened counterfactual thinking, and intensified feelings of regret. Their discussion of the concept of normality included experimental research illustrating various factors influencing the perception and interpretation of normality (Feldman & Albarracín, 2017; Feldman, 2018). This approach often employed past behavior as a reference point to assess the normality of an individual's actions.

@@

This study aims to directly replicate three experiments that demonstrated the exceptionality effect, representing early investigations into the influence of past behavior normality on regret. Specifically, we investigate hitchhiking and car accident scenarios from Kahneman and Miller (1986) in the first two part and replicate Miller and McFarland's (1986) Experiment 1, featuring a robbery scenario, in the part three to test norm theory. Although there is no documented direct replication of these experiments, the existing citation impact of Kahneman and Miller's work emphasizes the significance of a reexamination and detailed explanation of these effects. The study's findings are expected to contribute to a better understanding of the general nature of the exceptionality effect and the intricate influence of past behavior normality on regret.

>>>>

Regret, a pervasive and intense negative emotion, plays a pivotal role in shaping individuals' daily experiences. The recognition that other decisions or circumstances could have led to more favorable outcomes triggers regret (Byrne, 2016; Roese & Epstude, 2017; Kahneman & Tversky, 1982). Influential work by Kahneman and Miller (1986) in norm theory has highlighted the connection between abnormal behavior, heightened counterfactual thinking, and intensified feelings of regret. Their discussion of the concept of normality included experimental research illustrating various factors influencing the perception and interpretation of normality (Feldman & Albarracín, 2017; Feldman, 2018). This approach often employed past behavior as a reference point to assess the normality of an individual's actions.

@@

This study aims to directly replicate three experiments that demonstrated the exceptionality effect, representing early investigations into the influence of past behavior normality on regret. Specifically, we investigate hitchhiking and car accident scenarios from Kahneman and Miller (1986) in the first two part and replicate Miller and McFarland's (1986) Experiment 1, featuring a robbery scenario, in the third part to test norm theory. There are three research questions: for hitchhiking scenarios from Kahneman and Miller (1986), we investigated the proportion for perceived regret, injunctive social norms, descriptive social norms, and negative affect. For car accident scenario, we studied the proportion for perceived regret and luck. And for the scenario of robbery, we investigated victim compensation and regret (with exception conditions combined). Although there is no documented direct replication of these experiments, the existing citation impact of Kahneman and Miller's work emphasizes the significance of a reexamination and detailed explanation of these effects. The study's findings are expected to contribute to a better understanding of the general nature of the exceptionality effect and the intricate influence of past behavior normality on regret.
